### 2019-08-22 maciek.tex
Omega(kl)

  \maciek{cluster B is special?}
  \maciek{If DET splits a component, we issue intra-component requests until DET joins it}

@@

  The input $I_2$ consists of requests $(x^{(1)}, c(x^{(1)})), (x^{(2)}, c(x^{(2)})),\ldots$.
  The node $x^{(i)}$ is a node in the current configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
  After a request $(x^{(i)}, c(x^{(i)}))$, the node $x^{(i)}$ is a part of a non-singleton component and must be relocated to some other cluster (as its current cluster hosts a component $a_1, \ldots, a_{k-1}$).
  Then, DET places the node $x^{(i+1)}$ in place of $x^{(i)}$, and the next request concerns that node.
  % Initially, $x^{(1)} = a_k$, and the next node $x^{(2)}$ is the that DET locates in place of the

@@

  Now, we formally define $c(x^{(i)})$, and the ordering of nodes in the initial configuration.
  Consider a set of nodes from an initial configuration of a cluster $C$ (where $C \neq A, C \neq B$).
  By $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence $\{x^{(i)}\}$.
  At this point, we fix an arbitrary other node of $C$ and we label it $v^C_0$, and we set $c(v^C_i) = v^C_0$ for all $i$.
  For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence $\{x^{(i)}\}$.
  %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
  Some nodes of $C$ might never appear in the sequence $\{x^{(i)}\}$, and we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not.

>>>>

 We say that an online algorithm DET is component-respecting if it never reaches a configuration with a component split between clusters.
  %Recall that if DET resides in a configuration with a component split between clusters, we issue intra-component requests until
   \maciek{If DET splits a component, we issue intra-component requests until DET joins it}

@@

  The input $I_2$ consists of requests $(x^{(1)}, c(x^{(1)})), (x^{(2)}, c(x^{(2)})),\ldots$.
  The node $x^{(i)}$ is a node in the current configuration of DET that is co-located in one cluster with the component $a_1, \ldots, a_{k-1}$.
  After a request $(x^{(i)}, c(x^{(i)}))$, the node $x^{(i)}$ is a part of a non-singleton component and must be relocated to some other cluster (as its current cluster hosts a component $a_1, \ldots, a_{k-1}$).
  Then, DET places the node $x^{(i+1)}$ in place of $x^{(i)}$, and the next request concerns that node.
  % Initially, $x^{(1)} = a_k$, and the next node $x^{(2)}$ is the that DET locates in place of the
  The sequence continues as long as after serving the request, a perfect partition of the input graph exists.

@@

  Now, we formally define $c(x^{(i)})$, and the ordering of nodes in the initial configuration.
  Consider a set of nodes from an initial configuration of a cluster $C$ (where $C \neq A$).
  By $v^C_1$ we label the first node of a cluster $C$ that appears in the sequence $\{x^{(i)}\}$.
  At this point, we fix an arbitrary other node of $C$ and we label it $v^C_0$, and we set $c(v^C_i) = v^C_0$ for all $i$.
  For the following $k-2$ nodes of the cluster $C$, we label them $v^C_2, v^C_3, \ldots$ in the order of appearance in the sequence $\{x^{(i)}\}$.
  %For a node $v^C_p = x^{(j)}$, we set $y^{(j)} = v^C_0$.
  Some nodes of $C$ might never appear in the sequence $\{x^{(i)}\}$ (e.g. DET might terminate prematurely or might reside in a configuration with components split), and we number them arbitrarily, with the restriction that the nodes that do appear in the sequence have smaller number than the ones that do not.

  Now, we claim that given the input $I_1I_2$, every component-respecting DET 
  % (that does not terminate prematurely and never resides in a configuration with components split)
  must produce a sequence $\{x^{(i)}\}$ of length $L = k\cdot \ell - (k-1) - 2 - (\ell-1) - 1$.
  Consider a request $(x^{(i)}, c(x^{(i)})$.
  If in this configuration at least one singleton component is present (other than the components of $(x^{(i)}, c(x^{(i)})$), then serving this request is possible (i.e., the perfect partition exists), and a singleton node appears as $x^{(i+1)}$, and the sequence continues.
  Otherwise, as we have a component $a_1, \ldots, a_{k-1}$, no perfect partition exists.

  Hence, the length of a sequence $\{x^{(i)}\}$ depends on the number of singleton components.
  Before the arrival of the first request from $I_2$, the number of singleton components is $L + (\ell-1) + 1$, and after serving the last request, the number of singleton components is $1$.
  %Every node $x^{(i)}$ must be a singleton component, otherwise DET resides in a configuration with components split (and such DETs either leave such configuration eventually or issue an infinite cost).
  After each request, we join this singleton component with another component.
  This join decreases the number of singleton components by one or by two: if this is the join of two singletons $v^C_0, v^C_1$ (there are $\ell-1$ such joins: for all clusters but $A$), then the number of singletons decrease by $2$, and it decreases by $1$ otherwise.
  We conclude that the number of requests $\{x^{(i)}\}$ to decrease the number of singletons to $1$ is $L$.

  \medskip

  Now, we define the input $I_2'$.
